Estimation of Manning's Friction Coefficient using DART Buoys Data during \tohoku Tsunami

### 2013-12-24 paper/intro.tex
restructure tsunami section

\section{Introduction}

Tsunamis in the past decade have been responsible for some of the most  deadly
and costly natural disasters ever recorded. Coastal communities  have faced this
hazard by assessing the risk they pose, attempting  to make informed decisions
about the likelihood that such an event would  occur with cost to life and
property.  To do this, computational models of  tsunami events are often
employed to explore multiple different scenarios to predict the water elevation
caused by the tsunamis. The accurate prediction  of water elevations requires
accurate estimation of many model parameters that  are either measured directly,
defined empirically, or inferred from sparsely collected data. Unfortunately,
since tsunami events are relatively rare,  there is an abundance of uncertainty
in the input data for these computational  models ranging from effects of the
domain such as bathymetry and friction parameterizations,  to the earthquake
source.  In this article, we aim to estimate the Manning's $n$ friction
coefficient. Our approach quantifies the uncertainty in the predicted water
elevation  and employs an inverse modeling approach to estimate the Manning's
$n$ coefficient using observations of water elevation measured during an event.

@@

The method proposed here utilizes recent developments in uncertainty
quantification (UQ) methods that allow non-intrusive probing of the sensitivity
of realistic tsunami  models to uncertain parameters, inferring those parameters
from specific  observations.  In particular, we implement an inverse modeling
approach that relies on the Bayesian inference approach.  We will also rely on
polynomial chaos (PC) expansions to construct a faithful surrogate of the
response of the large and complex geophysical model simulations.  THe
construction of the surrogate enables us to efficiently implement the Bayesian
inference formalism as it dramatically reduces the cost of sampling the
posterior distribution.

@@

\alert{maybe an outline for this paper}

>>>>

\section{Introduction}
\label{sec:intro}
Tsunamis in the past decade have been responsible for some of the most  deadly
and costly natural disasters ever recorded. Coastal communities  have faced this
hazard by assessing the risk they pose, attempting  to make informed decisions
about the likelihood that such an event would  occur with cost to life and
property.  To do this, computational models of  tsunami events are often
employed to explore multiple different scenarios to predict the water elevation
caused by the tsunamis. The accurate prediction  of water elevations requires
accurate estimation of many model parameters that  are either measured directly,
defined empirically, or inferred from sparsely collected data. Unfortunately,
since tsunami events are relatively rare,  there is an abundance of uncertainty
in the input data for these computational  models ranging from effects of the
domain such as bathymetry and friction parameterizations,  to the earthquake
source.  In this article, we aim to estimate the Manning's $n$ friction
coefficient. Our approach quantifies the uncertainty in the predicted water
elevation  and employs an inverse modeling approach to estimate the Manning's
$n$ coefficient using observations of water elevation measured during an event.

@@

The method proposed here utilizes recent developments in uncertainty
quantification (UQ) methods that allow non-intrusive probing of the sensitivity
of realistic tsunami  models to uncertain parameters, inferring those parameters
from specific  observations.  In particular, we implement an inverse modeling
approach that relies on the Bayesian inference approach.  We will also rely on
Polynomial Chaos (PC) expansions to construct a faithful surrogate of the
response of the large and complex geophysical model simulations.  The
construction of the surrogate enables us to efficiently implement the Bayesian
inference formalism as it dramatically reduces the cost of sampling the
posterior distribution.

@@

The structure of this manuscript is as follows. First, we motivate the UQ problem 
by discussing the uncertainties of the drag parameters in Section~\ref{sec:manning}. 
Section~\ref{sec:simulation} summarizes the forward simulations setup
and the DART buoys observational data. 
Section~\ref{sec:formu} describes the two main numerical method employed:
Bayesian inference formalism that brings together observations 
and model results is briefly described in Section~\ref{sec:inference} 
and PC expansions that were used to build the surrogate model are described
in Section~\ref{sec:uqpce}. Section~\ref{sec:results} describes the main results in 
three subsections: first, we present evidence that the surrogate is a faithful 
representation of the \geoclaw model; second, we use the surrogate to explore the 
response surfaces of simulated water elevations, and to compute their sensitivity 
to the control parameters; and third, we present the posterior distribution obtained 
via MCMC sampling and contrast it to the prior distributions. 
A discussion of our findings and methodology is presented in Section~\ref{sec:conc}.

### 2013-12-24 paper/tsunami.tex
restructure tsunami section

\section{Manning's Friction Coefficient}
%We use the widely used .  The parameter $n$ is also allowed to vary in space determined by the initial water depth to study effects of deep-water, near-shore, and on-shore sensitivities.

@@

\section{Simulation Setup}

In this section we briefly describe the forward computational model used, \geoclaw, and the case study used in the following uncertainty analysis, the great \tohoku earthquake with observations from NOAA DART buoys.

@@

\subsection{The \tohoku Tsunami} \label{ssub:tohoku}

The \tohoku earthquake and tsunami of 2011 has been the subject of a number of studies due to the wealth of observational evidence and severity of the tsunami.  The earthquake itself had an estimated magnitude of $\text{M}_\text{w}$ 9.0 and caused massive devastation across Japan.  In the case study presented, the simulation is based on work done in \cite{MacInnes:2013cr} and modified to use the specified variable friction field.

\subsubsection{Source Model}

@@

Since the previous study found that the source model used did not impact the off-shore gauge stations, the ones used here from \cite{Ammon:2011dm}, was chose the source model that was commonly available.  Access to the finest bathymetry located near the Japanese coastline was also not included as it primarily effected only the shore observations.

\subsubsection{Bathymetry}
The bathymetry is shown in Figure~\ref{fig:setup} in addition to the topography
of the Pacific Ocean.
\subsubsection{Observations}

>>>>

\section{Manning's Friction Coefficient}
\label{sec:manning}
%We use the widely used .  The parameter $n$ is also allowed to vary in space determined by the initial water depth to study effects of deep-water, near-shore, and on-shore sensitivities.

@@

\section{The \tohoku Tsunami} 
\label{ssub:tohoku}

The \tohoku earthquake and tsunami of 2011 has been the subject of a number of studies due to the wealth of observational evidence and severity of the tsunami.  The earthquake itself had an estimated magnitude of $\text{M}_\text{w}$ 9.0 and caused massive devastation across Japan.  In the case study presented, the simulation is based on work done in \cite{MacInnes:2013cr} and modified to use the specified variable friction field.
In this section we briefly describe the forward computational model used, \geoclaw, and the 
simulation setup used in the uncertainty analysis and finally describe
the observations from NOAA DART buoys.

%\section{Simulation Setup}
%\label{sec:simulation}

@@

\subsection{Source Model and Bathymetry}

@@

Since the previous study found that the source model used did not impact the off-shore gauge stations, the ones used here from \cite{Ammon:2011dm}, was chose the source model that was commonly available.  Access to the finest bathymetry located near the Japanese coastline was also not included as it primarily effected only the shore observations.
\comment{The bathymetry is shown in Figure~\ref{fig:setup} in addition to the topography
of the Pacific Ocean.}

\subsection{DART Buoys Observations}
